\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2025-02-23 Manuscript/manuscript.tex
revised discussion paragraph 1

This study sought to answer two primary questions: were RTs of athletes who
competed at the 2022 World Track and Field Championships faster than
at other races, and is the 0.1-second RT barrier fair?
Regarding the first question, our analyses suggest that athletes
indeed had faster RTs at the 2022 World Championships compared to comparable
meets.  Our findings are strengthened by having multiple comparisons: athletes
who competed in 2019 and 2022 got older over that timespan but we included an
analysis in which athletes would have been younger in 2022 and then older in
2023.  There are validly raised concerns that athletes who compete at their
national level competitions in 2022 may not react at their fastest due to a less
competitive environment or because they would try to peak for a more important
championship.  However, the athletes who competed in 2019 and 2023 are not
impacted by this factor, so its effect is minimzed to just one subset of a
broader analysis.

Ideally, a centralized database containing RTs from
all World Athletics-certified meets would have allowed for a more
comprehensive analysis. Currently, however, RTs are seldom
included in the results available on the World Athletics website. This
limitation constrained our ability to evaluate athletes like Devon
Allen, as data from his races leading up to the World Championships in
2022 were not accessible. A more robust dataset would enable deeper
insights by allowing comparisons across a larger set of races.

In addressing the second question, our analyses of the GAMLSS model
reveal that the 0.1-second RT barrier, while rare, may not
be as extraordinary  as traditionally assumed. For men,
Table~\ref{tab:Sim_time} shows that RTs below 0.1 seconds
occur with a probability of approximately one in 362 starts  when
including the 2022 data. Lowering the threshold to 0.08 seconds
drastically  reduces this likelihood, supporting the idea that the
current barrier could be adjusted to reflect more realistic
probabilities of false starts. A similar pattern is observed for
women, as detailed in the Supplementary Material, where RTs
 below 0.1 seconds are exceedingly rare for the 100-meter dash
and 100-meter  hurdles. However, the uniformity of the 0.1-second
barrier for both men and women  is questionable, given numerous
studies documenting gender differences in reaction  times
\citep[e.g.,][]{lipps2011implications, babicc2009reaction,
  panoutsakopoulos2020gender}. These studies suggest that the current
threshold may  unfairly penalize men, for whom sub-0.1-second reaction
times are more probable. \citet{brosnan2017effects} advocate for
gender-specific barriers, a position that  aligns with our findings
and highlights the importance of tailoring thresholds to biological
distinctions.

>>>>

This study first examined whether reaction times (RTs) at the 2022
World Track and Field Championships were significantly faster than at
other competitions. Our analyses indicate that RTs at the 2022 World
Championships were consistently faster than those recorded at both
national-level competitions earlier in the same year and the 2019 and
2023 World Championships. The persistence of this pattern across
different comparison groups suggests that these differences are not
due to random variation or individual improvements over time. A more
comprehensive analysis would benefit from a centralized database
containing RTs from all World Athletics-certified meets, but such data
are not consistently available. However, by incorporating competitions
from multiple years (2019, 2022, and 2023), the analysis accounts for
potential confounding factors such as seasonality and age, as athletes
at different stages of their careers are represented in different
comparisons.

This study further assessed whether the 0.1-second RT threshold is a
fair standard for disqualification. Our analyses of the GAMLSS model
suggest that while RTs below 0.1 seconds are rare, they may not be
be as extraordinary as traditionally assumed. For men,
Table~\ref{tab:Sim_time} shows that RTs below 0.1 seconds
occur with a probability of approximately one in 362 starts when
including the 2022 data. Lowering the threshold to 0.08 seconds
drastically reduces this likelihood, supporting the idea that the
current barrier could be adjusted to reflect more realistic
probabilities of false starts. A similar pattern is observed for
women, as detailed in the Supplementary Material, where RTs
 below 0.1 seconds are exceedingly rare for the 100-meter dash
and 100-meter  hurdles. However, the uniformity of the 0.1-second
barrier for both men and women  is questionable, given numerous
studies documenting gender differences in reaction  times
\citep[e.g.,][]{lipps2011implications, babicc2009reaction,
  panoutsakopoulos2020gender}. These studies suggest that the current
threshold may  unfairly penalize men, for whom sub-0.1-second reaction
times are more probable. \citet{brosnan2017effects} advocate for
gender-specific barriers, a position that  aligns with our findings
and highlights the importance of tailoring thresholds to biological
distinctions.

### 2025-02-23 Manuscript/reply2.tex
revised discussion paragraph 1

\jy{Admit the limitation of the study in the discussion and point to
  the location of the addition.}

We have added additional explanations in the discussion section to address both
seasonality and age, although we don't think either one will be largely
impactful.  By including times from athletes in both 2019 and 2022, there are
athletes who got older from 2019 to 2022 but athletes in 2022 will be younger
than they were in 2023.  Including both comparisons is meaningful as it helps
diminish this effect.  Likewise while seasonality may impact the 2022 national
level comparison, the 2019 and 2023 comparisons mitigate this factor.

@@

We appreciate the thoughtful feedback and are glad the technical work
is recognized as solid. The revisions have focused on improving the
organization and clarity of the paper to better communicate the key
findings.

>>>>

We appreciate this thoughtful comment and have addressed it in the
discussion section. The revised text clarifies that comparisons across
multiple years (2019, 2022, and 2023) help mitigate the effects of
seasonality and age, as athletes contribute to different pairings
across years. Additionally, while seasonal peaking may influence RTs
at national competitions, the inclusion of both national and
international events provides a broader competitive context, reducing
the likelihood that seasonality alone explains the observed
differences. This clarification has been added to the first paragraph
of the discussion.

@@

We are delighted that the technical work
is recognized as solid. The revisions have focused on improving the
organization and clarity of the paper to better communicate the key
findings.
